Comment: Reasoning about Interference Between Units

### 2015-12-28 response.md

In the text we note that the SSR is particularly sensitive to differences in
means (or perhaps symmetric differences in tails that also produce a lot of
movement in the mean of one group relative to another group) while the KS-test
ought to have more power to detect asymmetric differences in the distributions
of the treated versus control groups --- for example, differences that change
variance without moving the means by stretching one tail of the distribution.

4) In letters we have an emphasis that new methods matter when applied to
real examples.  Can you demonstrate that a case where the SSR will change
the inferences we make when applied to a real world example?

We have added a section in which we replicate the analysis in Coppock's 2014
*Journal of Experimental Political Science* article "Information Spillovers:
Another Look at Experimental Estimates of Legislator Responsiveness".

# Response to Reviewer 1

We appreciate the positive review. In response to this review and the comments from the editor and other reviewer, we now valuate the SSR test statistic for three non-interference models (the sharp null model, the constant additive effects model, and the constant multiplicative effects model) and one new interference model (the one used by Coppock in his *Journal of Experimental Political Science* article).  We did not systematically search for other spillover models from political science, but agree that such a survey of models and now, test statistics, would be useful in the future.

# Response to Reviewer 2

We appreciate the positive review. We added a section in an attempt to provide more intuition about test statistics, models of effects, and their interaction. Continuing the 4 unit example did not, in the end, clarify Fisher/Rosenbaum-style hypothesis testing as we would have hoped so we only use that example in the beginning to illustrate the idea that each unit in a study may have many more than two potential outcomes. We did add a section showing a case where the KS test has more power than the SSR test -- in this case we used a skewed outcome with no spillover rather than heterogeneous exposure to spillover in order to keep things as simple as possible.

### 2015-12-29 response.md

Dear Professor Grimmer,

We really appreciate the constructive and timely feedback from you and the reviewers.

1) As Reviewer 1 suggests, how does the model perform as the assumptions
about spill overs deviate?  How sensitive are the inferences to the
particular assumptions you can make?

In the paper we clarified that we are focusing on testing sharp null
hypotheses arising from one model of how an experimental treatment might
propagate across a social network. We have added some discussion to the paper
highlighting the fact that the power of a test of a sharp null hypothesis
depends on *both* the model of causal effects *and* the test statistic (and
the effective sample size which itself depends on sample size and network
characteristics). We also note that, given randomization, any test statistic
ought to produce valid statistical inferences (i.e. tests that would reject
the truth no more than $100 \alpha \%$ of the time).

@@

4) In letters we have an emphasis that new methods matter when applied to
real examples.  Can you demonstrate that a case where the SSR will change
the inferences we make when applied to a real world example?

We have added a section in which we replicate the analysis in Coppock's 2014
*Journal of Experimental Political Science* article "Information Spillovers:
Another Look at Experimental Estimates of Legislator Responsiveness".

# Response to Reviewer 1

We appreciate the positive review. In response to this review and the comments from the editor and other reviewer, we now evaluate the SSR test statistic for three non-interference models (the sharp null model, the constant additive effects model, and the constant multiplicative effects model) and one new interference model (the one used by Coppock in his *Journal of Experimental Political Science* article).  We did not systematically search for other spillover models from political science, but agree that such a survey of models and now, test statistics, would be useful in the future.

### 2016-01-04 response.md

Dear Professor Grimmer,

We really appreciate the constructive and timely feedback from you and the
reviewers. Here is our response to your questions and suggestions and those of
the reviewers.

# Overall

@@

1) As Reviewer 1 suggests, how does the model perform as the assumptions
about spill overs deviate?  How sensitive are the inferences to the
particular assumptions you can make?

In the paper we clarified that we are focusing on testing sharp null
hypotheses arising from one model of how an experimental treatment might
propagate across a social network. We have added some discussion to the paper
highlighting the fact that the power of a test of a sharp null hypothesis
depends on *both* the model of causal effects *and* the test statistic (and
the effective sample size which itself depends on sample size and network
characteristics). We also note that, given randomization, any test statistic
ought to produce valid statistical inferences (i.e. tests that would reject
the truth no more than $100 \alpha \%$ of the time).

Since the point of the paper is to learn about test statistics, we do not systematically
explore alternative models of the propagation of causal effects across
networks. The original paper (BFP) on this method compared the performance of
tests when the data did and did not arise from the causal model and we point to
that paper for those who would like to pursue those set of questions. We did
add some discussion about the power of the SSR test statistic against different
kinds of simple causal effects and outcome distributions: the SSR test
statistic ought to be particularly sensitive to changes in mean differences
(after all, recall that the coefficient on a dichotomous variable in an least
squares model with a continuous outcome that minimizes the SSR is the
difference of means) and ought to be less sensitive to other differences than a
statistic like the KS test statistic.

@@

2) Reviewer 2 asks you to carry forward your toy example with four units
through the entire analysis.  If this isn't possible, please provide the
simplest case that illuminate the core intuition of your method.

When we set about adding the four unit example to the different sections of the
paper, we realized that we were not enhancing intuition: the two test
statistics applied to an experiment of size 4 mainly reflected the fact that we
had very little data. Instead, we added a discussion of the SSR and KS test
statistics applied to a case with no interference and with very simple models
of causal effect. We also added a figure and discussion that tries to show a bit more
about how models of effects work in the context of sharp null hypothesis
testing. We hope that this helps with the intuition so that the use of the SSR
test statistic in an experiment on a social network with a complex model of
treatment propagation is easier to understand.

### 2016-01-04 paper/introduction.tex
Fixing things to build the whole paper on keeling
Fixing up paper.
Edits pre resubmission

In a randomized experiment with $n=4$ subjects connected via a fixed network,
the response of subject $i=1$ might depend on the different ways that
treatment is assigned to the \emph{whole} network. When the treatment
assignment vector,$\bz$, provides treatment to persons 2 and 3,
$\bz=\{0,1,1,0\}$, person $i=1$ might respond one way,
$y_{i=1,\bz=\{0,1,1,0\}}$ and when treatment is assigned to persons 3 and 4,
$\bz=\{0,0,1,1\}$ person $i=1$ might act another way,
$y_{i=1,\bz=\{0,0,1,1\}}$. More generally, we might say that if the experiment had a causal effect on
person $i$, then her outcome would differ under different realizations of the
experimental treatment as a whole $y_{i,\bz} \ne y_{i,\bz'}$. The fundamental problem of
causal inference reminds us that we can never see both states of the world: we
only observe the outcome from person $i$ under one treatment assignment
vector, either $\bz$ or some $\bz'$ not both
\citep{holland:1986a,brady2008cae}.\footnote{In simpler settings, where
  treatment given to one individual has no effect on any other individual, we tend to
  write $y_{i,Z_i=1} \ne y_{i,Z_i=0}$ to say that treatment had a causal
  effect on person $i$.}

@@

Notice that the test statistic choice matters in this process: the engine of
statistical inference involves summarizing information against the
hypothesized claim, yet different ways to summarize information might be more
or less sensitive to substantively meaningful differences. The statistical
power of a simple test of the sharp null hypothesis of no effects will vary as
a function of the design of the study (total number of observations, proportion treated, blocking structure,
etc), characteristics of the outcome (continuous, binary, skewed, extreme
points, etc), and the way that a test statistic summarizes the outcome (does
it compare means, standard deviations, medians, qqplots, etc). In general,
test statistics should be powerful against relevant alternatives. \citet[\S 2.4.4]{rosenbaum:2002} provides more specific
advice about the large sample performance of certain classes of test statistics
and BFP repeat his general advice: ``Select a test statistic [$\mathcal{T}$] that will be
small when the treated and control distributions in the adjusted data \ldots
are similar, and large when the distributions diverge.'' \cite[Proposition 4
and 5, \S 2.9]{rosenbaum:2002} presents results proving that test statistics
with this property (``effect increasing'' test statistics), produce an
unbiased test of the hypothesis of no effects or positive effects when the
positive effects involve one parameter. Notice that this means that, a test of
the sharp null of no effects
based on a known randomization using \emph{any} effect increasing
test-statistic to be valid test (in that the test should produce $p$-values
less than $\alpha$ no more than 100$\alpha$ of the time when the null is true) even though different
test statistics may imply different power against false hypotheses.  Yet, when the models are complex and
may involve increasing effects in the direction of one parameter and
non-linear effects in the direction of another parameter, BFP showed that
sometimes a KS-test will have (1) no power to address the model at all such
that all hypothesized parameters would receive the same implausibility
assessment; (2) or might reject all such hypothesized parameters. Thus,
although in theory one may assess sharp multiparameter hypotheses,
 in practice one may not learn much from such tests. BFP thus recommended simulation studies of the operating
characteristics of tests as a piece of their workflow --- because  the theory
justifying simple one-dimensional effect increasing test statistics clearly
did not cover multi-parameter situations like those easily arising from social
network experiments.

@@

Fisher/Rosenbaum style randomization inference tends to use test statistics that
compare two distributions. Simple models imply that the distribution of the
outcome in the control remains fixed. For example, $\widetilde
y_{i,Z_i=0}=Y_i-Z_i \tau$ only changes the distribution of outcomes for units
in the treated condition. Comparing the mean of $\widetilde y_{i,Z_i=1}|Z_i=1$
to the mean of $\widetilde y_{i,Z_i=0}|Z_i=0$ makes intuitive sense in this
case, and, if $Y_i$ is Normal or at least unimodal without major outliers,
then this test might have optimal power. The complex model used as an example
by BFP involved adjustments to both control and treated outcomes --- some
hypothesized parameters would cause shifts in variance, others in location.
So, BFP proposed to use the KS-test statistic to assess the relationship
between $\widetilde y_{i,Z_i=0,Z_{-i}=0}$ and $Z_i$ (where $Z_{-i}=0$ means
``when all units other than $i$ are also not treated".)

Yet, one can also think about the process of hypothesis testing as a process
of assessing model fit, and there are usually better ways to evaluate the fit
of a model than comparing two marginal distributions: For example, the KS test
  uses the maximum difference in the empirical cumulative distributions of
  each treatment group calculated without regard for the relationship between
the treated and control distributions, thereby ignoring information that could
increase the precision of the test. The simplest version of the SSR test
statistic merely sums the difference between the mean of the outcome implied
by the hypothesis and individual outcomes:

@@

And we compared two models of effects --- a constant additive effects model in
which $y_{i,Z_i=1}=y_{i,Z_i=0}+\tau$ and a constant multiplicative effects
model in which $y_{i,Z_i=1}=y_{i,Z_i=0}*\tau$. We set the truth to be the
sharp null of no effects $\tau=0$ for the additive model and $\tau=1$ for the
multiplicative model. To further explain the process of hypothesis testing and
evaluation of the test statistics, we display the results from a power
analysis of one alternative hypothesis in figure~\ref{fig:boxplot}. A given
model and hypothesized parameter implies a distribution for observed outcomes
in the treatment group and control groups. Here, since we have no
interference, the hypotheses only have different implications for the control
group distribution. In this case, we chose a two parameters for the additive
and multiplicative models which produced very similar implications (as show by
the similarity in the ``Add. Model'' and ``Mult.  Model'' boxplots in both
panels of the figure). We see that when we tested the hypothesis of
$\tau_0=.34$ from the additive model using the KS test statistic on Normal
data, that we produced $p$-values lower than $\alpha=.05$ about 55\% of the
time across 1000 simulations. That is, the KS test has a power of about .55
to reject a false hypothesis with this Normal outcome and this constant
additive effects model. The analogous power for the SSR test statistic was
.73. For these particular parameters, we see the SSR performing better with
Normal outcomes but not being very effected by the model of effects (which
makes some sense here because we are choosing parameters at which both models
imply more or less the same patterns in outcomes).

\begin{figure}[H]\centering
  \includegraphics[width=.5\textwidth]{ksvsssr-boxplot.pdf}
  \caption{Distributions of the simulated data from left to right: $y_0$
    outcome with no experiment, observed outcome after random assignment,
    outcomes implied by a constant additive model of effects, outcomes implied by a constant
    multiplicative model. The hypothesized model parameters $\tau_0$ that
    produce the patterns shown are printed on the plot. The proportion of
    simulated tests in which these values of $\tau_0$ with $p$-values less
    than $\alpha=.05$ by the KS and SSR test statistics are printed on the plot as $\text{pow}_{KS}$ and
  $\text{pow}_\text{SSR}$. The SSR test statistic has more power for the
  Normal outcomes and less power for the skewed outcome.}\label{fig:boxplot}
\end{figure}

When we assessed power across many alternative hypotheses, our intuition was that the SSR
would have more power than the KS test when the outcome was Normal and when
the observational implication of the model of effects would shift means (i.e.\
the additive model). We used direct simulation of the randomization
distribution to generate $p$-values and repeated that process 1000 times to
gauge the proportion of rejections of false hypotheses (i.e.\ the power of the
tests). The results, not shown here but part of the reproduction archive for
this paper, bear out this intuition: the SSR has slightly more power than KS
for Normal outcomes in both the additive and multiplicative effects
conditions. SSR has slightly less power than KS when the outcome is skewed for
both models. In general, the SSR ought to be most powerful when the effect of
the experiment involves a shift in the location of the distributions of the
treated versus the controls. And it should be less powerful when the effect of
the experiment is mostly to leave the center of the distribution alone, but,
instead to stretch or compress the tails, or to concentrate the experimental
effect on one quantile.

>>>>

In a randomized experiment with $n=4$ subjects connected via a fixed network,
the response of subject $i=1$ might depend on the different ways that
treatment is assigned to the \emph{whole} network. When the treatment
assignment vector, $\bz$, provides treatment to persons 2 and 3,
$\bz=\{0,1,1,0\}$, person $i=1$ might respond one way,
$y_{i=1,\bz=\{0,1,1,0\}}$ and when treatment is assigned to persons 3 and 4,
$\bz=\{0,0,1,1\}$ person $i=1$ might act another way,
$y_{i=1,\bz=\{0,0,1,1\}}$. More generally, we might say that if the experiment had a causal effect on
person $i$, then her outcome would differ under different realizations of the
experimental treatment as a whole $y_{i,\bz} \ne y_{i,\bz'}$. The fundamental problem of
causal inference reminds us that we can never see both states of the world: we
only observe the outcome from person $i$ under one treatment assignment
vector, either $\bz$ or some $\bz'$ not both
\citep{holland:1986a,brady2008cae}.\footnote{In simpler settings, where
  treatment given to one individual has no effect on any other individual, we tend to
  write $y_{i,Z_i=1} \ne y_{i,Z_i=0}$ to say that treatment had a causal
  effect on person $i$.}

@@

Notice that the test statistic choice matters in this process: the engine of
statistical inference involves summarizing information against the
hypothesized claim, yet different ways to summarize information might be more
or less sensitive to substantively meaningful differences. The statistical
power of a simple test of the sharp null hypothesis of no effects will vary as
a function of the design of the study (total number of observations, proportion treated, blocking structure,
etc), characteristics of the outcome (continuous, binary, skewed, extreme
points, etc), and the way that a test statistic summarizes the outcome (does
it compare means, standard deviations, medians, qqplots, etc).\footnote{Some
  use the term `effective sample size' --- which we first saw in
  \citet{kish65} ---
  to highlight the fact that statistical power depends on more than the number
  of rows in a given rectangular dataset. } In general,
test statistics should be powerful against relevant alternatives. \citet[\S 2.4.4]{rosenbaum:2002} provides more specific
advice about the large sample performance of certain classes of test statistics
and BFP repeat his general advice: ``Select a test statistic [$\mathcal{T}$] that will be
small when the treated and control distributions in the adjusted data \ldots
are similar, and large when the distributions diverge.'' \cite[Proposition 4
and 5, \S 2.9]{rosenbaum:2002} presents results proving that test statistics
with this property (``effect increasing'' test statistics), produce an
unbiased test of the hypothesis of no effects or positive effects when the
positive effects involve one parameter. Such results mean that a test of
the sharp null of no effects
based on a known randomization using \emph{any} effect increasing
test-statistic to be valid test (in that the test should produce $p$-values
less than $\alpha$ no more than 100$\alpha \%$ of the time when the null is true) even though different
test statistics may imply different power against false hypotheses.  Yet, when the models are complex and
may involve increasing effects in the direction of one parameter and
non-linear effects in the direction of another parameter, BFP showed that
sometimes a KS-test will have (1) no power to address the model at all such
that all hypothesized parameters would receive the same implausibility
assessment; (2) or might reject all such hypothesized parameters. Thus,
although in theory one may assess sharp multiparameter hypotheses,
 in practice one may not learn much from such tests. BFP thus recommended simulation studies of the operating
characteristics of tests as a piece of their workflow --- because  the theory
justifying simple one-dimensional effect increasing test statistics clearly
did not cover multi-parameter situations like those easily arising from social
network experiments.

@@

Fisher/Rosenbaum style randomization inference tends to use test statistics that
compare two distributions. Simple models imply that the distribution of the
outcome in the control remains fixed. For example, $\widetilde
y_{i,Z_i=0}=Y_i-Z_i \tau$ only changes the distribution of outcomes for units
in the treated condition. Comparing the mean of $\widetilde y_{i,Z_i=1}|Z_i=1$
to the mean of $\widetilde y_{i,Z_i=0}|Z_i=0$ makes intuitive sense in this
case, and, if $Y_i$ is Normal or at least unimodal without major outliers,
then this test using means as test statistics might have optimal power. The complex model used as an example
by BFP involved adjustments to both control and treated outcomes --- some
hypothesized parameters would cause shifts in variance, others in location.
So, BFP proposed to use the KS-test statistic to assess the relationship
between $\widetilde y_{i,Z_i=0,Z_{-i}=0}$ and $Z_i$ (where $Z_{-i}=0$ means
``when all units other than $i$ are also not treated".)

Yet, one can also think about the process of hypothesis testing as a process
of assessing model fit, and there are usually better ways to evaluate the fit
of a model than comparing two marginal distributions: For example, the KS test
  uses the maximum difference in the empirical cumulative distributions of
  each treatment group calculated without regard for the relationship between
the treated and control distributions, thereby ignoring information that could
increase the precision of the test. The simplest version of the SSR test
statistic merely sums the difference between the mean of the outcome implied
by the hypothesis and individual outcomes thereby including the correlation
between treated and control outcomes as a part of the distribution of the
statistic:

@@

We compared two models of effects --- a constant additive effects model in
which $y_{i,Z_i=1}=y_{i,Z_i=0}+\tau$ and a constant multiplicative effects
model in which $y_{i,Z_i=1}=y_{i,Z_i=0} \cdot \tau$. We set the truth to be
the sharp null of no effects such that the true $\tau=0$ for the additive
model and the true $\tau=1$ for the multiplicative model. To further explain
the process of hypothesis testing and evaluation of the test statistics, we
display the results from a power analysis of one alternative hypothesis in
figure~\ref{fig:boxplot}. A given model and hypothesized parameter implies a
distribution for observed outcomes in the treatment group and control groups.
Here, since we have no interference, the hypotheses only have different
implications for the control group distribution. In this case, we chose a two
parameters for the additive and multiplicative models which produced very
similar implications (as show by the similarity in the ``Add. Model'' and
``Mult.  Model'' boxplots in both panels of the figure). We see that when we
tested the hypothesis of $\tau_0=0.34$ from the additive model using the KS
test statistic on Normal data, that we produced $p$-values lower than
$\alpha=0.05$ about 55\% of the time across 1000 simulations. That is, the KS
test has a power of about 0.55 to reject a false hypothesis with this Normal
outcome and this constant additive effects model. The analogous power for the
SSR test statistic was 0.73. For these particular parameters, we see the SSR
performing better with Normal outcomes but not being very effected by the
model of effects (which makes some sense here because we are choosing
parameters at which both models imply more or less the same patterns in
outcomes). On non-Normal outcomes (shown in the "Jittered Geometric Outcomes"
panel), the SSR test statistic had less power. Again, the particular model of
effect did not matter in this case because we chose the alternative hypothesis
($\tau_0$ on the plots) to represent the case where both models implied
similar outcome distributions.

\begin{figure}[H]\centering
  \includegraphics[width=.9\textwidth]{ksvsssr-boxplots.pdf}
  \caption{Distributions of the simulated data from left to right: $y_0$
    outcome with no experiment, observed outcome after random assignment,
    outcomes implied by a constant additive model of effects ("Add. Model"), outcomes implied by a constant
    multiplicative model ("Mult. Model"). The hypothesized model parameters $\tau_0$ that
    produce the patterns shown are printed on the plot. The proportion of
    simulated tests in which these values of $\tau_0$ with $p$-values less
    than $\alpha=.05$ by the KS and SSR test statistics are printed on the plot as $\text{pow}_{KS}$ and
  $\text{pow}_\text{SSR}$. The SSR test statistic has more power for the
  Normal outcomes and less power for the skewed outcome.}\label{fig:boxplot}
\end{figure}

When we assessed power across many alternative hypotheses, our intuition was
that the SSR would have more power than the KS test when the outcome was
Normal and when the observational implication of the model of effects would
shift means (i.e.\ the additive model). We used direct simulation of the
randomization distribution to generate $p$-values and repeated that process
1000 times to gauge the proportion of rejections of a range of false
hypotheses  (i.e.\ the power of the tests at many different values of
$\tau_0$). The results, not shown here but part of the reproduction archive
for this paper, bear out this intuition: the SSR has slightly more power than
KS for Normal outcomes in both the additive and multiplicative effects
conditions. SSR has slightly less power than KS when the outcome is skewed for
both models. In general, the SSR ought to be most powerful when the effect of
the experiment involves a shift in the location of the distributions of the
treated versus the controls. And it should be less powerful when the effect of
the experiment is mostly to leave the center of the distribution alone, but,
instead to stretch or compress the tails, or to concentrate the experimental
effect on one quantile.
